GMSE: an R package for generalised management strategy evaluation

### 2018-05-04 notebook/ms/SI5.Rmd
Text changes throughout
Changes relate mostly to the framing of the conflict situation. Also some minor changes throughout.

D'sfl;As part of the [ConFooBio](https://sti-cs.org/confoobio/) project, a central focus of GMSE is on simulating the management of populations of conservation interest, with a particular emphasis on understanding conservation conflict; further development of GMSE is expected to continue with this as a priority, further building upon the decision-making algorithms of managers and users to better understand how conflict arises and can potentially be resolved. Hence, GMSE is not intended as a substitute for packages such as [FLR](http://www.flr-project.org/), but the integration of these packages with GMSE could make use of GSME's current and future simulation capabilities, and particularly the [genetic algorithm](). Such integration might be possible using the `gmse_apply` function, which allows for [custom defined submodels]() to be used within the GMSE framework, and with default GMSE submodels. Hence, GMSE might be especially useful for modelling the management of fisheries under conditions of increasing harvesting demands and stakeholder conflict. We do not attempt such an ambitious project here, but instead show how such a project could be developed through integration of [FLR](http://www.flr-project.org/) and `gmse_apply`.

Here we follow a [Modelling Stock-Recruitment with FLSR](http://www.flr-project.org/doc/Modelling_stock_recruitment_with_FLSR.html) example, then integrate this example with `gmse_apply` to explore the behaviour of simulated fishers who are goal-driven to maximise their own harvest. We emphasise that this example is provided only as demonstration of how GMSE can potentially be integrated with already developed fisheries models, and is not intended to make recommendations for management in any population.

@@

```{r}
newFL   <- FLSR(); # Initialises the empty FLSR object
data(nsher_data);
```

@@

An initial run of these models gives values of `r round(new_rec, digits = 2)` for `new_rec` and `r round(obs_rec, digits = 2)` for `obs_rec`. We are now ready to use the built-in manager and user submodels in `gmse_apply`. We will assume that managers attempt to keep a recruitment of 5000, and that there are 4 independent fishers who attempt to maximise their catch. We assign a user budget of `manager_budget = 10000`, and all other values are set to GMSE defaults. In the built-in GMSE functions, the manager will use the estimate of recruitment based on `obs_rec` and use it to set the cost of harvesting (`culling` in GMSE).

@@

The above output from `sim_sum` reports the recruitment (resource or operational model), recruitment estimate (observation error model), management (harvest control model), and user (implementation model) simulation results. This example simulation demonstrates the ability of GMSE to integrate with fisheries libraries such as [FLR](http://www.flr-project.org/) through `gmse_apply`. In addition to being a useful wrapping function for MSE submodels, `gmse_apply` can therefore be used to take advantage of the genetic algorithm in the GMSE default manager and user models. This flexibility will be retained in future versions of `gmse_apply`, allowing custom resource and observation models that are built for specific systems to be integrated with an increasingly complex genetic algorithm simulating various aspects of human decision-making.

>>>>

As part of the [ConFooBio](https://sti-cs.org/confoobio/) project, a central focus of GMSE is on simulating the management of animal populations of conservation interest, with a particular emphasis on understanding conservation conflict; further development of GMSE is expected to continue with this as a priority, further building upon the decision-making algorithms of managers and users to better understand how conflict arises and can be managed and mitigated. Hence, GMSE is not intended as a substitute for packages such as [FLR](http://www.flr-project.org/), but the integration of these packages with GMSE could make use of GSME's current and future simulation capabilities, and particularly the [genetic algorithm](). Such integration might be possible using the `gmse_apply` function, which allows for [custom defined submodels]() to be used within the GMSE framework, and with default GMSE submodels. Hence, GMSE might be especially useful for modelling the management of fisheries under conditions of increasing competing staleholder demands and conflicts. We do not attempt such an ambitious project here, but instead show how such a project could be developed through integration of [FLR](http://www.flr-project.org/) and `gmse_apply`.

Here we follow a [Modelling Stock-Recruitment with FLSR](http://www.flr-project.org/doc/Modelling_stock_recruitment_with_FLSR.html) example, then integrate this example with `gmse_apply` to explore the behaviour of a number of simulated fishers who are goal-driven to maximise their own harvest and a manager that aims to keep the fish stocks at a predefined target level. The core concept in GMSE is that manager can only incentivise fishers to harvest less or more by varying the cost of fishing (through e.g. taxes) given a set manager budget; please note that the manager cannot force the fisher to follow any policy. Based on the cost of fishing, the fisher can then given their own budget decide whether to invest in fishing or keep the budget. This concepts represents a nartural resource managmeent and conservation conflict, where one party aims to maximise their livelihood (fisher) and the other aims to keep a population at a sustainable level and prevent it from going extinct. Importantly, the manager does not have full control over fishers but can set policies to incentivise sustainable behaviour. We emphasise that this example is provided only as demonstration of how GMSE can potentially be integrated with already developed fisheries models, and is not intended to make recommendations for management in any population.

@@

```{r}
newFL   <- FLSR(); # Initialises the empty FLSR object
data(nsher_data); #called from GMSE library rather than nsher from FLCore library
```

@@

An initial run of these models gives values of `r round(new_rec, digits = 2)` for `new_rec` and `r round(obs_rec, digits = 2)` for `obs_rec`. We are now ready to use the built-in manager and user submodels in `gmse_apply`. We will assume that managers attempt to keep a recruitment of 5000, and that there are 4 independent fishers [stakehodlers in gmse_apply says 10, is that a difference?] who attempt to maximise their catch. We assign a user budget of `manager_budget = 10000`, and all other values are set to GMSE defaults. In the built-in GMSE functions, the manager will use the estimate of recruitment based on `obs_rec` and use it to set the cost of harvesting (`culling` in GMSE).

@@

The above output from `sim_sum` reports the recruitment (resource or operational model), recruitment estimate (observation error model), management set harvest cost (harvest control model), user harvested numbers (implementation model) and spawning stock biomass (SSB) simulation results. This example simulation demonstrates the ability of GMSE to integrate with fisheries libraries such as [FLR](http://www.flr-project.org/) through `gmse_apply`. In addition to being a useful wrapping function for MSE submodels, `gmse_apply` can therefore be used to take advantage of the genetic algorithm in the GMSE default manager and user models. This flexibility will be retained in future versions of `gmse_apply`, allowing custom resource and observation models that are built for specific systems to be integrated with an increasingly complex genetic algorithm simulating various aspects of human decision-making.
